---
title: "Scaling_Function_Demo"
output: html_document
date: "2023-11-01"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This document explains the usage of the min_max_scaling function defined in the R script file within this repository.

The function is tested on the USArrests data, a builtin R dataset on US crime statistics from 1974. 
```{r, echo=FALSE,suppresswarn}
library(tidyverse)
#US Arrests - dataset of crime statistics from 1974 in USA
data <- USArrests
```


The function is coded as follows:
```{r cars}
#function returns scaled value for each crime statistic observation, where max value is defined as input by user, each observations is transformed into its proportion of the max value 
min_max_scaling <- function(data,max_value){
  scaled_data <- data %>%
    select(Murder, Assault, Rape) %>% 
    mutate(Murder_Scaled = ((Murder - min(Murder))/(max(Murder)-min(Murder)))*max_value) %>% 
    mutate(Assault_Scaled = ((Assault - min(Assault))/(max(Assault)-min(Assault)))*max_value) %>% 
    mutate(Rape_Scaled = ((Rape - min(Rape))/(max(Rape)-min(Rape)))*max_value) %>% 
    select(Murder_Scaled,Rape_Scaled,Assault_Scaled)
  return(scaled_data)
}
```

As seen, the function takes the arguments 'data' and 'max_value.' This second argument 'max_value' defines the maximum of the scaled data the function returns. 

To demonstrate this, we can test the function for different max values, both large and small. Choosing different values is useful for relative comparison of crime across the USA.
```{r}
#max_value = 100
min_max_scaling(data,100)
```

As seen above, choosing 100 as the max value transforms each observation into its percentage of highest observed crime rate. For instance, Vermont has just 8.43% the murder rate of the highest state, Georgia, which is 100%.

The function also works for values less than 1, if that is what we want our maximum to be for data-analysis purposes. 
```{r}
#max_value = 0.5
min_max_scaling(data,0.5)
```








